Positive Correlation Between Racialization, Gendering and Police Strip Searches in the Toronto Municipal, 2020-2021

### 2024-01-26 outputs/paper/paper.qmd
finalize paper for submission
retitle graphs, tables and finalize paper

```{r}
#| echo: false
#| message: false
#| label: tbl-cleaned_race_gender_data
#| tbl-cap: Sample of Race and Gender Data against Police Arrests with Strip Searches

@@

```{r fig.pos="H"}
#| echo: false
#| message: false
#| label: tbl-cleaned_search_reasons_data
#| tbl-cap: Sample of Reasons for Police Strip Searches Data

@@

```{r fig.pos="H"}
#| echo: false
#| message: false
#| label: tbl-gender-strip-search
#| tbl-cap: Summary Statistics of Gendered Police Strip Searches

@@

```{r fig.width=7, fig.height=3, fig.pos="H"}
#| echo: false
#| warning: false
#| message: false
#| label: fig-reasons-race
#| fig-cap: Relationship Between Reasons for Strip Search and Race in Toronto, 2020-2021
#| layout-ncol: 1
#| fig-subcap: ["Black and White","Other Minority Groups and White"]

@@

```{r fig.height=7, fig.width=9, fig.pos='H'}
#| echo: false
#| message: false
#| label: fig-race-gender-search
#| fig-cap: Racial Identities and Police Strip Search Per Arrest in Toronto, 2020-2021
cleaned_male_female_data <-
  cleaned_race_gender_data |>
  filter(gender != "U")

@@

Overall, the data unfolds the relationship between strip searching and social identities such as race and gender. Within these intersecting identities, it was found that if there is an police arrest in Toronto from 2020 to 2021, strip searching would most likely happen to an Indigenous person or a black person. This is a statistically improbable insight, considering the major population make-up of white people in Toronto and their huge counts of arrests as seen in past bar graphs (see [@fig-race-gender-search]). Out of all the minority racial groups, at 2434 people, the black population has the highest counts of strip searches. Regarding reasons accounted for each strip search, minority racial groups are more likely to be strip searched without any reason provided or with less reasons provided compared to those that are ethnically white. Gender-wise, even though arrested males prevail in terms of being strip searched (see [@fig-race-gender-search]), females are more likely to be searched with less reasons provided. These insights all substantiate the assumptions that Toronto police strip searches are racialized and gendered, which calls for a legal and political revisions of this intrusive form of policing as it hinges on racial biases, profiling, social intimidation and sexual invasion.

@@

While these findings point towards a positive correlation between police strip search being racially inclined and gendered, particularly towards the black and female population of Toronto, it's worth noting how the data has decontextualized the reasons the police account for each strip search. Specifically, each reason-- including "Injury", "Weapons", "Escape" and "Possess Evidence"--is quantified as a single, equal unit for each arrest, without other further details on the severity scale of the reason itself. This can very much seclude the fact that some arrests--even those having the same reason--are more necessary to be strip searched than others, thus skewing the nature of this paper's analysis. Moreover, referring back to @tbl-reasons-strip-search, more reasons to account for a strip search doesn't necessarily mean that a person is bound to be strip searched. The number of reasons the dataset attributes to each strip search--as provided by Toronto Police Service [@torontoPolice]--can be purely projected based on the police in power's racial preconceptions or gender biases. This means that the way the reasons are flagged out for each strip search has might be personally motivated.

>>>>

```{r}
#| echo: false
#| message: false
#| label: tbl-cleaned_race_gender_data
#| tbl-cap: Sample of Race and Gender against Police Arrests with and without Strip Searches

@@

```{r fig.pos="H"}
#| echo: false
#| message: false
#| label: tbl-cleaned_search_reasons_data
#| tbl-cap: Sample of Reasons for Police Strip Searches

@@

```{r fig.pos="H"}
#| echo: false
#| message: false
#| label: tbl-gender-strip-search
#| tbl-cap: Summary Statistics of Gender Identities against Police Strip Searches

@@

```{r fig.width=7, fig.height=2.5, fig.pos="H"}
#| echo: false
#| warning: false
#| message: false
#| label: fig-reasons-race
#| fig-cap: Relationship Between Reasons for Strip Search and Race in Toronto, 2020-2021
#| layout-ncol: 1
#| fig-subcap: ["Black and White","Other Minority Groups and White"]

@@

```{r fig.height=7, fig.width=9, fig.pos='H'}
#| echo: false
#| message: false
#| label: fig-race-gender-search
#| fig-cap: Count of Gender and Racial Identities against Police Strip Searches in Toronto
cleaned_male_female_data <-
  cleaned_race_gender_data |>
  filter(gender != "U")

@@

Illustrated in @fig-race-gender-search, this paper aims to interogate the relationship between strip searching and social identities such as race and gender in Toronto from 2020 to 2021. Within these intersecting identities, it was found that if there is an police arrest in Toronto from 2020 to 2021, strip searching would most likely happen to an Indigenous person or a black person. This is a statistically improbable insight, considering the major population make-up of white people in Toronto and their huge counts of arrests as seen in past bar graphs (see [@fig-race-gender-search]). Out of all the minority racial groups, at 2434 people, the black population has the highest counts of strip searches. Regarding reasons accounted for each strip search, minority racial groups are more likely to be strip searched without any reason provided or with less reasons provided compared to those that are ethnically white. Gender-wise, even though arrested males prevail in terms of being strip searched (see [@fig-race-gender-search]), females are more likely to be searched with less reasons provided. These insights all substantiate the assumptions that Toronto police strip searches are racialized and gendered, which calls for a legal and political revisions of this intrusive form of policing as it hinges on racial biases, profiling, social intimidation and sexual invasion.

@@

While these findings point towards a positive correlation between police strip search being racially inclined and gendered, particularly towards the black and female population of Toronto, it's worth noting how the data has decontextualized the reasons the police account for each strip search. Specifically, each reason-- including "Injury", "Weapons", "Escape" and "Possess Evidence"--is quantified as a single, equal unit for each arrest, without other further details on the severity scale of the reason itself. This can very much seclude the fact that some arrests--even those having the same reason--are more necessary to be strip searched than others, thus skewing the nature of this paper's analysis. Moreover, referring back to @tbl-reasons-strip-search, more reasons to account for a strip search doesn't necessarily mean that a person is bound to be strip searched. The number of reasons the dataset attributes to each strip search--as provided by Toronto Police Service [@torontoPolice]--can be purely projected based on the police in power's racial preconceptions or gender biases. This means that the way the reasons are flagged out for each strip search might be personally motivated.
